Theories of deliberated preference

### 2018-02-27 simple.tex
First part done (simple theories with decisive arguments).

\section{Decision situation}
The object of study in a given decision situation is the deliberated preferences of a given individual $i$, about a topic $\alltopic$, given a set of arguments $\allargs$, given a certain query protocol, and over a certain time frame.

A decision situation is defined as a triple $(\alltopic, \allargs, \ileadstost)$. It concerns a specific individual $i$, whose deliberated preferences are studied, and whose reactions to arguments is modeled by $\ileadstost$, as will be defined shortly.

The decision situation contains a topic denoted by $\alltopic$, which is a set of propositions $\prop \in \alltopic$ about which we are interested of knowing the deliberated preferences of $i$. Propositions are not described further, but are supposedly understandable by $i$ (for example, they could be sentences in some natural language). Alternatively, in the case of an analyst helping $i$ to make a decision, $\alltopic$ is a set of propositions about which $i$ is interested to know his own deliberated preferences. I assume $\alltopic$ is closed under negation and introduce a symbol $¬$ for negating a proposition: if $\prop \in \alltopic$, then $¬\prop \in \alltopic$, and $¬(¬\prop) = \prop$.

@@

\begin{example}
	Consider a decision problem where a set of alternatives $\allalts$ is given, food products among which $i$ would like to form a deliberated preference (taking into account the effects of food on health, price, pleasure, morality issues related to the production process, and so on). The topic could be defined as $\alltopic$ = $\Set{\prop_\alt, ¬\prop_\alt, \forall \alt \in \allalts}$, where $\prop_\alt$ is the proposition according to which $\alt$ is one of the best alternatives among $\allalts$ (there is no strictly better alternatives among $\allalts$, from $i$’s point of view), and $¬\prop_\alt$ is the proposition according to which $\alt$ is not one of the best alternatives among $\allalts$ (some $\alt' \in \allalts$ is a strictly better alternative, from $i$’s point of view). 
	The set $\allargs$ is defined as the set of all strings, because in this example it is considered acceptable to restrict arguments to those that can take a textual form.
\end{example}

As vNM aptly summarize, “It is clear that every measurement – or rather every claim of measurability – must ultimately be based on some immediate sensation, which possibly cannot and certainly need not be analyzed any futher.” The fundamental element here is the reaction of $i$ to arguments. Given a proposition $\prop$ and two arguments $\ar, \ar' \in \allargs$, a query consists in presenting both arguments to $i$ and observing which proposition $i$ considers valid in her current state of mind, thus using both arguments and possibly other arguments she has in mind: is it that $\prop$ holds, that $¬\prop$ holds, or that both are possible? The third possibility allows for the case where no argument appears more decisive than the other one.

@@

This information is only indirectly observable: for example, if $(\ar, \ar') \ileadsto^\prop \{\prop\}$, a querier will never know more than $(\ar, \ar') \ileadsto^\prop C$ with $\prop \in C \subseteq \Set{\prop, ¬\prop, \text{both}}$. This limitation of our observation will not be a problem for the results of this article. Intuitively, this bears on the fact that if a theory claims that $(\ar, \ar') \ileadsto^\prop \{\prop\}$, then either this claim is correct, or it is falsifiable, in the sense that there is a possibility of observing a contradiction to its claim.

I voluntarily leave the querying protocol not precisely specified, as multiple reasonable choices are possible and the results of this article depend only on specific assumptions that will be made clear shortly. One possible choice, when arguments and propositions are strings, is to read both arguments to $i$, together with $\prop$, and ask $i$ to verbally report the choice he opts for among the three possibilities. Another is to present both arguments and ask $i$ to pick an item among some choice set and make the choice correspond to a validation of $\prop$ or $¬\prop$, possibly even letting the choice engage $i$, for example telling $i$ that she may keep the object of choice (this is illustrated in \cref{ex:pick}). This second possibility is especially interesting as it makes the protocol partially observable in the restricted sense usually appreciated in revealed preferences approaches. Note that depending on the protocol, the choice “both” is not necessarily observable (more about this below).
%In that case, the semantics of $\prop$ and $¬\prop$ relatedly change: instead of meaning that $\prop$ is definitely the 
%Indeed, \citet{tversky_intransitivity_1969} note that individuals “are not perfectly consistent in their choices. When faced with repeated choices between x and y, people often choose x in some instances and y in others. Furthermore, such inconsistencies are observed even in the absence of systematic changes in the decision maker’s taste which might be due to learning or sequential effects. It seems, therefore, that the observed inconsistencies reflect inherent variability or momentary fluctuation in the evaluative process.” This is an accepted fact of experimental psychology \citep{luce_utility_2000}. However, it is reasonable to suspect that fluctuations occur less often, if at all, for some choices. For example, facing a choice between tea or coffee, I would virtually never pick tea (all other things being equal). We are interested in capturing precisely those kind of preferences.

@@

To summarize this information, define $\Phi = \alltopic × \set{\text{possible}, \text{sure}}$, and define a single relation $\ileadstost \subseteq (\allargs × \allargs) × \Phi$, pronounced “always leads to”. Elements in $\Phi$ will also be called propositions, which should create no ambiguity. Define $\forall \phi = (\prop, \text{possible}) \in \Phi: (\ar, \ar') \ileadstost \phi ⇔ (\ar, \ar') \ileadsto^\prop C$ with $C \subseteq \Set{\prop, \text{both}}$; and $\forall \phi = (\prop, \text{sure}) \in \Phi: (\ar, \ar') \ileadstost \phi ⇔ (\ar, \ar') \ileadsto^\prop \set{\prop}$. 
The semantics are that whenever $(\ar, \ar') \ileadstost (\prop, \text{sure})$, $i$ always considers that $\prop$ is sure (meaning that $¬\prop$ is excluded), when presented with these two arguments, and whenever $(\ar, \ar') \ileadstost (\prop, \text{possible})$, $i$ always considers that $\prop$ is “at least” possible (meaning that $¬\prop$ may be possible as well), when presented with these two arguments.

Define the negation operator $¬$ over $\Phi$ as: $¬(\prop, \text{possible}) = (¬\prop, \text{sure})$ and $¬(\prop, \text{sure}) = (¬\prop, \text{possible})$. It follows that $¬¬\phi = \phi$. Write $(\ar, \ar') \nileadstost \phi$ for $¬[(\ar, \ar') \ileadstost \phi]$ and define $(\ar, \ar') \ileadstoe \phi$ as equivalent to $(\ar, \ar') \nileadstost ¬\phi$. 
The relation $\ileadstoe$ is pronounced “sometimes leads to”.
A proposition $\phi$ is about $\prop$ iff $\phi \in \{\prop, ¬\prop\} × \{\text{possible}, \text{sure}\}$. 
\footnote{\label{ft:sttoe}From (A1) and (A2) we can write, using this notation, $(\ar, \ar') \ileadstost \phi ⇒ (\ar, \ar') \ileadstoe \phi$.}

Actually we will use only the $\ileadstopst$ and $\ileadstosst$ ternary relations from now on (and their negations $\ileadstose$ and $\ileadstope$), together with the property (A1) that $\ileadstopst \subseteq \ileadstope$
\footnote{Equivalently: $\ileadstosst \subseteq \ileadstose$, $(\ar, \ar') \ileadstose ¬\prop ∨ (\ar, \ar') \ileadstope \prop$, $\ileadstopst \prop ⇒ ¬ (\ileadstosst ¬\prop)$, $\ileadstosst ¬\prop ⇒ ¬ (\ileadstopst \prop)$.} 
and (A2) that $\ileadstosst \subseteq \ileadstopst$
\footnote{Equivalently: $\ileadstose \subseteq \ileadstope$.}, 
the pair $(\ar, \ar')$ being understood as unordered (thus $(\ar, \ar') \ileadstosst \prop ⇔ (\ar', \ar) \ileadstosst \prop$, and similarly for $\ileadstopst$).

When “both” is not observable, the same definitions hold, with $f_t(\ar, \ar') ≠ b$. It follows that $\ileadstopst = \ileadstosst$. We say in this case that $\ileadstost$ does not distinguish sure from possible. Thus, elements $\phi \in \Phi$ may be considered as simply equal to $\prop$ or $¬\prop$, as $(\prop, \text{sure})$ is treated exactly as $(\prop, \text{possible})$.

\begin{definition}[Decision situation]
	A decision situation is a topic $\Phi$, a set of arguments $\allargs$, an arbiter relation $\ileadstost \subseteq (\allargs × \allargs) × \Phi$.
\end{definition}

$\ar$ is $\phi\text{-dec} ⇔ \forall \ar': (\ar', \ar) \ileadstost \phi$. 
\footnote{Equivalently: $\ar \ \phidec ⇔ \exists \ar \suchthat [\ileadstostinv(\phi)](\ar) = \allargs$.}
\footnote{We also say that $\ar$ is $\propsuredec$, $\proppossdec$, or $¬\proppossdec$, when $\phi=(\prop, \text{sure})$, $\phi=(\prop, \text{possible})$, or $\phi=(¬\prop, \text{possible})$, respectively.
$\propsuredec$ ⇔ $\forall \ar': (\ar', \ar) \ileadstosst \prop$, $\proppossdec$ ⇔ $\forall \ar': (\ar', \ar) \ileadstopst \prop$.}

The following proposition follows from the fact that when $\ar^+$ is $\phidec$, given any $\ar^−$, $(\ar^+, \ar^−) \ileadstost \phi$ hence $(\ar^+, \ar^−) \ileadstoe \phi$ (using \cref{ft:sttoe}).
\begin{proposition}
	\label{thm:nocontrdec}
	$\exists \ar \ \phidec ⇒ \nexists \ar' \ ¬\phidec.$
\end{proposition}

The deliberated judgment of $i$ maps any unordered pair $(\prop, ¬\prop)$ to one of four possibilities. For easier treatment, we summarize $i$’s deliberated judgment in a single set $\iPhi \subseteq \Phi$, as follows. (Interpretation is given in parenthesis.)
\begin{itemize}
	\item $(\prop, \text{possible}) \in \iPhi ∧ (\prop, \text{sure}) \in \iPhi ∧ (¬\prop, \text{possible}) \notin \iPhi ∧ (¬\prop, \text{sure}) \notin \iPhi$ iff $\exists \ar \ \propsuredec$ ($\prop$ is sure)
	\item $(\prop, \text{possible}) \in \iPhi ∧ (¬\prop, \text{possible}) \in \iPhi ∧ (\prop, \text{sure}) \notin \iPhi ∧ (¬\prop, \text{sure}) \notin \iPhi$ iff $\exists \ar \ \proppossdec$ and $\exists \ar \ ¬\proppossdec$ (both $\prop$ and $¬\prop$ are possible)
	\item $\phi \notin \iPhi$ for all $\phi$ about $\prop$ iff neither $\prop$ is sure, nor $¬\prop$ is sure, nor both are possible ($(\prop, ¬\prop)$ is undecided).
\end{itemize}
\Cref{thm:nocontrdec} shows that the second case is separate from the first.
\footnote{Example of $(\prop, ¬\prop)\text{-undecided}$: $\forall (\ar, \ar'), (\ar, \ar') \ileadstose ¬\prop, (\ar, \ar') \ileadstose \prop$.}
\footnote{Example of $\overline{¬\prop\text{-CC}} ∧ \prop \in \iposs$ (thus with $\prop\text{-CC}$ and no $(\prop, ¬\prop)$-unstability): $\forall (\ar, \ar'), (\ar, \ar') \ileadstose \prop, (\ar, \ar') \ileadstope ¬\prop, (\ar, \ar') \ileadstopst \prop$. In such a situation, it is unclear whether $\prop$ holds rather than $¬\prop$ or both $\prop$ and $¬\prop$.}
\footnote{Something like Axiom JU should be sufficient for CC. We say $(\ar, \ar')$ is $\prop$-unstable iff $(\ar, \ar') \ileadstose \prop ∧ (\ar, \ar') \ileadstose ¬\prop$. JU: If $(\ar, \ar')$ is $\prop$-unstable, then $\exists \ar_2 \in \allargs$ such that for $(\ar_1, \ar_0) = (\ar, \ar')$ or for $(\ar_1, \ar_0) = (\ar', \ar)$, $\ar_2$ attacks decisively $\ar_1$, where we mean by this that $\ar_2$ combined with $\ar_0$ is decisive. Means that if not CC, then not JU. Hence, infinite discussion happens, or irreducible unstability. Or, the memory buffer of $i$ is too short to hold adequate reinstated arguments, in which case our approach based on decisive arguments is not adequate.}

\begin{proposition}
	a) $\phi \in \iPhi ⇒$ some $\ar$ is $\phidec$. b) $(\prop, \text{possible}) \in \iPhi$ iff $\exists \ar \ \proppossdec ∧ (\exists \ar \ \propsuredec ∨ \exists \ar \ ¬\proppossdec)$. c) $(\prop, \text{sure}) \in \iPhi ⇒ (\prop, \text{possible}) \in \iPhi$. d) $\phi \in \iPhi ⇒ ¬\phi \notin \iPhi$. e) $(\prop, ¬\prop)$ undecided $⇔ \phi \notin \iPhi ∧ ¬\phi \notin \iPhi$ for some $\phi$ about $\prop$. f) If $\phi \in \iPhi$ for some $\phi$ about $\prop$, then $\phi \in \iPhi ⇔ ¬\phi \notin \iPhi$ for all $\phi$ about $\prop$. g) If $\ileadstost$ does not distinguish sure from possible, no pair of proposition is mapped to “both are possible”.
\end{proposition}
\begin{proof}
	a) is a consequence of A2. b) and c) follow from the definition of $\iPhi$. d) is obtained using a), \cref{thm:nocontrdec}, and the contrapositive of a). About e), we must show that $\phi \notin \iPhi ∧ ¬\phi \notin \iPhi$ for some $\phi$ about $\prop$ implies $\phi' \notin \iPhi$ for all $\phi'$ about $\prop$. Assume $\phi=(l, \text{possible})$ and $¬\phi = (¬l, \text{sure})$ are not in $\iPhi$, with $l = t$ or $l = ¬t$. We need to prove that $(l, \text{sure})$ and $(¬l, \text{possible})$ are not in $\iPhi$ either. We obtain $(l, \text{sure}) \notin \iPhi$ using the contrapositive of c). And $(¬l, \text{possible}) \notin \iPhi$ because $(¬l, \text{possible}) \in \iPhi$ would require another $\phi'$ about $\prop$ to be present in $\iPhi$, as can be checked from the definition of $\iPhi$. f) follows from d) together with the fact that $(\prop, ¬\prop)$ is (by definition) not undecided when $\phi \in \iPhi$ for some $\phi$ about $\prop$, hence, from e), $\phi \in \iPhi ∨ ¬\phi \in \iPhi$. Finally, g) is because in that case, $\exists \ar \ \proppossdec ⇔ \exists \ar \ \propsuredec$, thus using \cref{thm:nocontrdec}, $\exists \ar \ \proppossdec ⇒ \nexists \ar' \ ¬\proppossdec$.
\end{proof}

\section{Theories proposing decisive arguments}
In most situations, there are too many arguments to exhaustively test whether any $\ar$ is a decisive argument for any $\phi$. Furthermore, it is impossible to confirm definitely that $(\ar, \ar') \ileadstost \phi$, whenever $(\ar, \ar') \ileadstost \phi$. We rather propose to introduce theories, and expose them to falsification.

A theory $T$ proposing decisive arguments describes a set $D_T$ of decision situations to which it applies, and proposes, for any situation $d = (\Phi, \allargs, \ileadstost) \in D_T$, a subset $\Phi_{d, T} \subseteq \Phi$ and a mapping $f_{d, T}: \Phi_{d, T} → \allargs$. The theory claims that it knows the deliberated judgment of $i$ (the individual described by $\ileadstost$ in $d$) about the propositions in $\Phi_{d, T}$, and claims that $f_{d, T}(\phi)$ is a decisive argument for $\phi$.

>>>>

\section{Decision situation}
The object of study in a given decision situation is the deliberated preferences of a given individual $i$. 
A decision situation concerns a topic $\alltopic$, given a set of arguments $\allargs$, given a certain query protocol, and over a certain time frame. The relation $\ileadstost$, to be defined shortly, models the reaction of $i$ to arguments about the topic. The decision situation will be defined as a triple $(\alltopic, \allargs, \ileadstost)$ satisfying some conditions to be defined after having presented those three fundamental elements.

The decision situation relates to a topic denoted by $\alltopic$, which is a set of propositions $\prop \in \alltopic$ about which we are interested of knowing the deliberated preferences of $i$. Propositions are not described further, but are supposedly understandable by $i$ (for example, they could be sentences in some natural language). Alternatively, in the case of an analyst helping $i$ to make a decision, $\alltopic$ is a set of propositions about which $i$ is interested to know his own deliberated preferences. I assume $\alltopic$ is closed under negation and introduce a symbol $¬$ for negating a proposition: if $\prop \in \alltopic$, then $¬\prop \in \alltopic$, and $¬(¬\prop) = \prop$.

@@

\begin{example}
	Consider a decision problem where a set of alternatives $\allalts$ is given, containing food products among which $i$ would like to form a deliberated preference (taking into account the effects of food on health, price, pleasure, morality issues related to the production process, and so on). The topic could be defined as $\alltopic$ = $\Set{\prop_\alt, ¬\prop_\alt, \forall \alt \in \allalts}$, where $\prop_\alt$ is the proposition according to which $\alt$ is one of the best alternatives among $\allalts$ (there is no strictly better alternatives among $\allalts$, from $i$’s point of view), and $¬\prop_\alt$ is the proposition according to which $\alt$ is not one of the best alternatives among $\allalts$ (some $\alt' \in \allalts$ is a strictly better alternative, from $i$’s point of view). 
	The set $\allargs$ is defined as the set of all strings, because in this example it is considered acceptable to restrict arguments to those that can take a textual form.
\end{example}

Define $\Phi = \alltopic × \set{\text{possible}, \text{sure}}$. Elements in $\Phi$ will also be called propositions, which should create no ambiguity. 
The relation $\ileadstost \subseteq (\allargs × \allargs) × \Phi$, pronounced “always leads to”, has the following semantics.
Whenever $(\ar, \ar') \ileadstost (\prop, \text{sure})$, $i$ always considers that $\prop$ is sure (meaning that $¬\prop$ is excluded), when presented with these two arguments, and whenever $(\ar, \ar') \ileadstost (\prop, \text{possible})$, $i$ always considers that $\prop$ is “at least” possible (meaning that $¬\prop$ may be possible as well), when presented with these two arguments. In both cases, the term “always” means that the consideration is stable over time, including after having presented other arguments to $i$. 
Two query protocols presented below will illustrate how these semantics might be satisfied. 
However, I voluntarily leave the querying protocol not precisely specified in general, as multiple reasonable choices are possible. The results of this article depend only on assumptions about $\ileadstost$.

Define the negation operator $¬$ over $\Phi$ as: $¬(\prop, \text{possible}) = (¬\prop, \text{sure})$ and $¬(\prop, \text{sure}) = (¬\prop, \text{possible})$. It follows that $¬¬\phi = \phi$. Write $(\ar, \ar') \nileadstost \phi$ for $¬[(\ar, \ar') \ileadstost \phi]$ and define $(\ar, \ar') \ileadstoe \phi$ as equivalent to $(\ar, \ar') \nileadstost ¬\phi$. 
The relation $\ileadstoe$ is pronounced “sometimes leads to”.
A proposition $\phi$ is about $\prop$ iff $\phi \in \{\prop, ¬\prop\} × \{\text{possible}, \text{sure}\}$.

The pair $(\ar, \ar')$ is to be understood as unordered (thus $(\ar, \ar') \ileadstost \prop ⇔ (\ar', \ar) \ileadstost \prop$).

The relation $\ileadstost$ must satisfy the two following properties, or axioms. (A1) $\ileadstost \subseteq \ileadstoe$
\footnote{Equivalently: $\ileadstosst \subseteq \ileadstose$, $\ileadstopst \subseteq \ileadstope$, $(\ar, \ar') \ileadstose ¬\prop ∨ (\ar, \ar') \ileadstope \prop$, $\ileadstopst \prop ⇒ ¬ (\ileadstosst ¬\prop)$, $\ileadstosst ¬\prop ⇒ ¬ (\ileadstopst \prop)$.} 
and (A2) $\forall \prop \in \alltopic: (\ar, \ar') \ileadstost (\prop, \text{sure}) ⇒ (\ar, \ar') \ileadstost (\prop, \text{possible})$
\footnote{Equivalently: $\ileadstose \subseteq \ileadstope$.}. 
Observe that (A1) can equivalently be formulated as $\forall \prop \in \alltopic: [(\ar, \ar') \ileadstost (\prop, \text{sure}) ⇒ (\ar, \ar') \ileadstoe (\prop, \text{sure})]$, or equivalently, $\forall \prop \in \alltopic: [(\ar, \ar') \ileadstost (\prop, \text{possible}) ⇒ (\ar, \ar') \ileadstoe (\prop, \text{possible})]$, and (A2) is equivalent to $\forall \prop \in \alltopic: [(\ar, \ar') \ileadstoe (\prop, \text{sure}) ⇒ (\ar, \ar') \ileadstoe (\prop, \text{possible})]$.

\begin{definition}[Decision situation]
	A decision situation is a topic $\alltopic$ closed under negation, a set of arguments $\allargs$, an “always leads to” relation $\ileadstost \subseteq (\allargs × \allargs) × \Phi$ satisfying (A1) and (A2) and using unordered pairs of arguments, $\Phi$ being the set of propositions $\alltopic × \set{\text{sure}, \text{possible}}$.
\end{definition}

\subsection{Query protocols}
As vNM aptly summarize, “It is clear that every measurement – or rather every claim of measurability – must ultimately be based on some immediate sensation, which possibly cannot and certainly need not be analyzed any futher.” The fundamental element here is the reaction of $i$ to arguments, which will serve as the informational basis to define $\ileadstost$. Given a proposition $\prop$ and two arguments $\ar, \ar' \in \allargs$, a query consists in presenting both arguments to $i$ and observing which proposition $i$ considers valid in her current state of mind, thus using both arguments and possibly other arguments she has in mind: is it that $\prop$ holds, that $¬\prop$ holds, or that both are possible? The third possibility allows for the case where no argument appears more decisive than the other one.

@@

This information is only indirectly observable: for example, if $(\ar, \ar') \ileadsto^\prop \{\prop\}$, a querier will never know more than $(\ar, \ar') \ileadsto^\prop C$ with $\prop \in C \subseteq \Set{\prop, ¬\prop, \text{both}}$. 
Also, it may be only possible to test one ordering of the queries (assuming that $i$ forgets over time might permit to test different orderings, but this assumption might be unrealistic). In some cases, it may even be only possible to test a single query with a given individual.
This limitation of our observation will not be a problem for the results of this article. Intuitively, this bears on the fact that, first, theories apply to multiple individuals, and second, if a theory claims that $(\ar, \ar') \ileadsto^\prop \{\prop\}$, then either this claim is correct, or it is falsifiable, in the sense that there is a possibility of observing a contradiction to its claim.

To be more concrete, one possible choice of a query protocol, when arguments and propositions are strings, is to read both arguments to $i$, together with $\prop$, and ask $i$ to verbally report the choice he opts for among the three possibilities. Another is to present both arguments and ask $i$ to pick an item among some choice set and make the choice correspond to a validation of $\prop$ or $¬\prop$, possibly even letting the choice engage $i$, for example telling $i$ that she may keep the object of choice (this is illustrated in \cref{ex:pick}). This second possibility is especially interesting as it makes the protocol partially observable in the restricted sense usually appreciated in revealed preferences approaches. Note that depending on the protocol, the choice “both” is not necessarily observable (more about this below).
%In that case, the semantics of $\prop$ and $¬\prop$ relatedly change: instead of meaning that $\prop$ is definitely the 
%Indeed, \citet{tversky_intransitivity_1969} note that individuals “are not perfectly consistent in their choices. When faced with repeated choices between x and y, people often choose x in some instances and y in others. Furthermore, such inconsistencies are observed even in the absence of systematic changes in the decision maker’s taste which might be due to learning or sequential effects. It seems, therefore, that the observed inconsistencies reflect inherent variability or momentary fluctuation in the evaluative process.” This is an accepted fact of experimental psychology \citep{luce_utility_2000}. However, it is reasonable to suspect that fluctuations occur less often, if at all, for some choices. For example, facing a choice between tea or coffee, I would virtually never pick tea (all other things being equal). We are interested in capturing precisely those kind of preferences.

@@

Define $\ileadstost$ from $\ileadsto^\prop$ as follows, for any $\prop \in \alltopic$: $(\ar, \ar') \ileadstost (\prop, \text{possible}) ⇔ (\ar, \ar') \ileadsto^\prop C$ with $C \subseteq \Set{\prop, \text{both}}$; and $(\ar, \ar') \ileadstost (\prop, \text{sure}) ⇔ (\ar, \ar') \ileadsto^\prop \set{\prop}$.

Note that (A1) and (A2) hold.

When “both” is not observable, $\ileadstost$ is defined as: $(\ar, \ar') \ileadstost (\prop, \text{possible}) ⇔ (\ar, \ar') \ileadstost (\prop, \text{sure}) ⇔ (\ar, \ar') \ileadsto^\prop \set{\prop}$. (A1) and (A2) still hold. We say in this case that $\ileadstost$ does not distinguish sure from possible. Thus, elements $\phi \in \Phi$ may be considered as simply equal to $\prop$ or $¬\prop$, as $(\prop, \text{sure})$ is treated exactly as $(\prop, \text{possible})$.

\subsection{Deliberated preference}
The following important definition is required in order to define the deliberated preference of $i$. An argument $\ar$ is decisive for $\phi$, in short, $\phi\text{-dec}$ iff $\forall \ar' \in \allargs: (\ar', \ar) \ileadstost \phi$. 
\footnote{Equivalently: $\ar \ \phidec ⇔ [\ileadstostinv(\phi)](\ar) = \allargs$, and $\exists \ar \ \phidec ⇔ [\ileadstostinv(\phi)]^{-1}(\allargs) ≠ \emptyset$.}
I also say that $\ar$ is $\propsuredec$, $\proppossdec$, or $¬\proppossdec$, when $\phi=(\prop, \text{sure})$, $\phi=(\prop, \text{possible})$, or $\phi=(¬\prop, \text{possible})$, respectively.
\footnote{$\propsuredec$ ⇔ $\forall \ar': (\ar', \ar) \ileadstosst \prop$, $\proppossdec$ ⇔ $\forall \ar': (\ar', \ar) \ileadstopst \prop$.}

\begin{proposition}
	\label{thm:nocontrdec}
	For any decision situation, a) $\exists \ar \ \phidec ⇒ \nexists \ar' \ ¬\phidec$ and b) $\exists \ar \ \propsuredec ⇒ \nexists \ar' \ ¬\propsuredec$.
\end{proposition}
\begin{proof}
	About a), when $\ar$ is $\phidec$, given any $\ar'$, $(\ar, \ar') \ileadstost \phi$ hence $(\ar, \ar') \ileadstoe \phi$ (using (A1)), hence $(\ar, \ar') \nileadstost ¬\phi$. b) obtains similarly using (A2).
\end{proof}

The deliberated preference of $i$ maps each pair $(\prop, ¬\prop)$ to one among four possibilities, depending on the existence of decisive arguments supporting the propositions about $\prop$. The mapping is defined as follows.
\begin{description}
	\item[$\prop$ is sure] iff $\exists \ar \ \propsuredec$
	\item[$¬\prop$ is sure] iff $\exists \ar \ ¬\propsuredec$
	\item[both $\prop$ and $¬\prop$ are possible] iff $\exists \ar \ \proppossdec$ and $\exists \ar \ ¬\proppossdec$
	\item[$(\prop, ¬\prop)$ is undecided] iff neither $\prop$ is sure, nor $¬\prop$ is sure, nor both are possible.
\end{description}
\Cref{thm:nocontrdec} shows that the cases are separate.
\footnote{Example of $(\prop, ¬\prop)\text{-undecided}$: $\forall (\ar, \ar'), (\ar, \ar') \ileadstoe (¬\prop, \text{sure}), (\ar, \ar') \ileadstoe (\prop, \text{sure})$.}
\footnote{Example of $\overline{¬\prop\text{-CC}} ∧ \prop \in \iposs$ (thus with $\prop\text{-CC}$ and no $(\prop, ¬\prop)$-unstability): $\forall (\ar, \ar'), (\ar, \ar') \ileadstose \prop, (\ar, \ar') \ileadstope ¬\prop, (\ar, \ar') \ileadstopst \prop$. In such a situation, it is unclear whether $\prop$ holds rather than $¬\prop$ or both $\prop$ and $¬\prop$.}
\footnote{Something like Axiom JU should be sufficient for CC. We say $(\ar, \ar')$ is $\prop$-unstable iff $(\ar, \ar') \ileadstose \prop ∧ (\ar, \ar') \ileadstose ¬\prop$. JU: If $(\ar, \ar')$ is $\prop$-unstable, then $\exists \ar_2 \in \allargs$ such that for $(\ar_1, \ar_0) = (\ar, \ar')$ or for $(\ar_1, \ar_0) = (\ar', \ar)$, $\ar_2$ attacks decisively $\ar_1$, where we mean by this that $\ar_2$ combined with $\ar_0$ is decisive. Means that if not CC, then not JU. Hence, infinite discussion happens, or irreducible unstability. Or, the memory buffer of $i$ is too short to hold adequate reinstated arguments, in which case our approach based on decisive arguments is not adequate.}

It should be clear that the validity of the conclusions obtained from any theory of deliberated preference rests on the acceptance of this just given normative definition, namely, it is to be considered meaningful that the deliberated preference of $i$ be defined in such a way, given the query protocol used for defining $\ileadstost$ in a particular application.

This information is summarized by defining a subset $\iPhi \subseteq \Phi$ of propositions that represent $i$’s deliberated preference, defined as follows.
\begin{description}
	\item[$\prop$ is sure] ⇒ $(\prop, \text{possible}) \in \iPhi ∧ (\prop, \text{sure}) \in \iPhi ∧ (¬\prop, \text{possible}) \notin \iPhi ∧ (¬\prop, \text{sure}) \notin \iPhi$
	\item[both $\prop$ and $¬\prop$ are possible] ⇒ $(\prop, \text{possible}) \in \iPhi ∧ (¬\prop, \text{possible}) \in \iPhi ∧ (\prop, \text{sure}) \notin \iPhi ∧ (¬\prop, \text{sure}) \notin \iPhi$
	\item[$(\prop, ¬\prop)$ is undecided] ⇒ $\phi \notin \iPhi$ for all $\phi$ about $\prop$.
\end{description}
Because this defines unequivocally the membership of every proposition about $\prop$ for every $\prop \in \alltopic$, this suffices to define $\iPhi$; and it follows that the reverse implications hold as well.

The following propositions can help get convinced that the proposed definition of deliberated preference makes sense.
\begin{proposition}
	a) $\phi \in \iPhi ⇒$ some $\ar$ is $\phidec$. b) $(\prop, \text{possible}) \in \iPhi$ iff $\exists \ar \ \proppossdec ∧ (\exists \ar \ \propsuredec ∨ \exists \ar \ ¬\proppossdec)$. c) $(\prop, \text{sure}) \in \iPhi ⇒ (\prop, \text{possible}) \in \iPhi$. d) $\phi \in \iPhi ⇒ ¬\phi \notin \iPhi$. e) $(\prop, ¬\prop)$ undecided $⇔ \phi \notin \iPhi ∧ ¬\phi \notin \iPhi$ for some $\phi$ about $\prop$. f) If $\phi \in \iPhi$ for some $\phi$ about $\prop$, then $\phi \in \iPhi ⇔ ¬\phi \notin \iPhi$ for all $\phi$ about $\prop$. g) If $\ileadstost$ does not distinguish sure from possible, no pair of proposition is mapped to “both are possible”.
\end{proposition}
\begin{proof}
	a) is a consequence of A2. b) and c) follow from the definition of $\iPhi$. d) is obtained using a), \cref{thm:nocontrdec}, and the contrapositive of a). About e), we must show that $\phi \notin \iPhi ∧ ¬\phi \notin \iPhi$ for some $\phi$ about $\prop$ implies $\phi' \notin \iPhi$ for all $\phi'$ about $\prop$. Assume $\phi=(l, \text{possible})$ and $¬\phi = (¬l, \text{sure})$ are not in $\iPhi$, with $l = t$ or $l = ¬t$. We need to prove that $(l, \text{sure})$ and $(¬l, \text{possible})$ are not in $\iPhi$ either. We obtain $(l, \text{sure}) \notin \iPhi$ using the contrapositive of c). And $(¬l, \text{possible}) \notin \iPhi$ because $(¬l, \text{possible}) \in \iPhi$ would require another $\phi'$ about $\prop$ to be present in $\iPhi$, as can be checked from the definition of $\iPhi$. f) follows from d) together with the fact that $(\prop, ¬\prop)$ is (by definition) not undecided when $\phi \in \iPhi$ for some $\phi$ about $\prop$, hence, from e), $\phi \in \iPhi ∨ ¬\phi \in \iPhi$. Finally, g) is because in that case, $\exists \ar \ \proppossdec ⇔ \exists \ar \ \propsuredec$, thus using \cref{thm:nocontrdec}, $\exists \ar \ \proppossdec ⇒ \nexists \ar' \ ¬\proppossdec$.
\end{proof}

It is important to remark that when the query protocol does not admit observation of “both”, thus, when $\ileadstost$ does not distinguish sure from possible, it does \emph{not} follow that $i$ is sure of either $\prop$ or $¬\prop$ for every pair of proposition: the “undecided” case will hold whenever $i$ considers no argument as sufficiently strong to determine a decisive preference (assuming that in that case, $i$ at least sometimes alternates his choice). Admittedly, it might also happen that $i$ is really indifferent between $\prop$ and $¬\prop$ but still systematically chooses, say, $\prop$ over $¬\prop$, in which case it will be (somewhat) erroneously concluded that $\prop \in \iPhi$. This is the price to pay for using protocols that do not allow to observe indifference. Observe however that (depending on the application) this error might be considered not harmful, as the obtained model will be faithful to $i$’s behavior after deliberation.

\section{Theories proposing decisive arguments}
In most situations, there are too many arguments to attempt to exhaustively test whether any $\ar$ is a decisive argument for any $\phi$. And most importantly, it is impossible to confirm definitely that $(\ar, \ar') \ileadstost \phi$, whenever $(\ar, \ar') \ileadstost \phi$. This is why falsifiable theories are needed.

A theory $T$ proposing decisive arguments describes a set $D_T$ of decision situations to which it applies, and proposes, for any situation $d = (\Phi, \allargs, \ileadstost) \in D_T$, a subset $\Phi_{d, T} \subseteq \Phi$ and a mapping $f_{d, T}: \Phi_{d, T} → \allargs$. The theory claims that it knows the deliberated preference of $i$ (the individual described by $\ileadstost$ in $d$) about the propositions in $\Phi_{d, T}$, and claims that $f_{d, T}(\phi)$ is a decisive argument for $\phi$. The theory $T$ is said to have a position about $\prop$ if $(\prop, \text{sure}) \in \Phi_{d, T} ∨ (¬\prop, \text{possible}) \in \Phi_{d, T}$. 
In order for the claims of the theory to be interpretable, it is requested that if $T$ has a position about $\prop$, then it must have a position about $¬\prop$. This forces $T$, for each pair $(\prop, ¬\prop)$, to either claim that one is sure, claim that both are possible, or have no claim about the pair.
